\coloc: Optimize Spatial Visualization Following Gestalt Principles

### 2023-03-17 Manuscript/manuscript.tex
revise use-cases.

\section*{Use Case 1: Spatially Differential Colocalization in DLPFC}
To demonstrate the new visualization, we apply \coloc to visualize the colocalization of two genes, (FYN) and (EFNA5) in the human dorsolateral prefrontal cortex. The two genes are previously identified that relates to SCZ expression via cell-cell communication analysis. The purpose of the visualization is to see if the density of colocalization are different across different cortex layers. Using \coloc, we can simultaneously display the colocalization and cortex layer in the same spatial plot, instead of having two saprate plots, one for cortex layer and one for colocalization, making the verification of the visual within the spaital context more straightforward and easier.

To create a colocalization plots, \coloc follows a similar principle as \texttt{ggplot2}'s layering strategy. In the first step, we will use the funciton \texttt{spCOCOON} to create a base plot that converts a spatial transcroptimcs data structure to a plot object. In a second step, we use the function \texttt{add\_fill} to color code each spots with colcalization status, which includes three levels, spots that expressed both genes and spots that expressed either of the genes. Then, we will apply the function \texttt{add\_COCOON} to show the spatial domain (cortex layer) information. Optionally, one can apply the funciton \texttt{add\_symbol} to display another layer of information. Here, we use  the symbol to highlight the colocalization spots, i.e. the spots that expresses both genes. Because we followed the layering strategy that \texttt{ggplot2}. One can seamsely to integrate \coloc to any \texttt{ggplot2}-based spatial visiulization by piping a \texttt{ggplot2} object to \texttt{add\_COCOON} and \texttt{add\_symbol} and customize the plots using the \texttt{ggplot2} syntex and themes, greatly reduce the techinical burden to customize these new spatial plot designs.

\section*{Use Case 2: Multi-dimensional UMAP}

>>>>

\section*{Use Case 1: Spatially Differential Colocalization in DLPFC}
In a recent study investigating the molecular organization of human dorsolateral prefrontal cortex (DLPFC) \cite{huukimyers_2023}, two schizophrenia risk genes, membrane-bound ligand ephrin A5 (\textit{EFNA5}) and ephrin type-A receptor 5 (\textit{EPHA5}), are identified to colocalize via the cell-cell communication analysis. In addition, data suggest Layer 6 is the most highly co-localized layer compared to other cortex layers. To visually examine the inference, we apply \coloc to create a multi-dimensional \textit{in-situ} spatial map (Figure \label{fig:visual}) that simultaneously exhibits the cortex layers (displayed with color-coded spot outlines) and the categorized colocalization status of \textit{FYN} and \textit{EFNA5} (displayed with color-coded spot fill). Compared to the traditional visualization where the cortex layers and the colocalization status are ploted in two side-by-side figures, our proposed visualization enables directly mapping colocalization status to the spatial domain, simplifying the perception of two sources of information and allowing cognitive comparison across cortex layers.

\section*{Use Case 2: Multi-dimensional UMAP}
The application of the proposed framework doesn't limit to in-situ visualizations of spatially resolved data. It can be broadly applied to data mapped to any 2-dimensional coordinate system to simultaneously display multiple variables. Such systems include euclidean space (including spatial coordinate as a special case) and data-driven embedding space, for example, UMAP and t-SNE. To demonstrate, we apply the proposed visualization to address the challenge of simultaneously displaying cluster membership and gene expression in a single-cell UMAP plot. To address the overplotting problem, \texttt{schex} proposed to apply hexagonal binning strategy to unbiasedly display the gene expression.\cite{freytag_2020} In addition, color-coded convex hulls were used to annotate different clusters of cells (Figure \label{fig:visual}). Nevertheless, the convex hulls create substantial overlapping areas, creating confusion when interpreting cluster memberships of hexagons in the overlapping areas. To improve the interpretability of the visualization, we integrate the proposed visualization to the \texttt{schex} by replacing convex hulls with color-coded hexagons boundaries to avoid possible membership confusion. The visual improvement is easily implemented without any modification of \texttt{schex} thanks to the Grammar of Graphics \cite{wilkinson_2012} standard.
